Notes on the Godunov Step in {\tt MAESTRO}

### 2008-04-15 docs/Godunov_notes/paper.tex
cleanup utrans section

For Cartesian problems, for the radial faces, we must include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$:
\begin{eqnarray}
\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib-\eb_r}^n+w_{0,r-1})\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{r-1},\label{Radial utrans_L} \\
\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}(\wt_{\ib}^n+w_{0,r})\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_r.\label{Radial utrans_R}
\end{eqnarray}
Note that $w_0$ is edge-centered, so we define:
\begin{equation}
w_{0,r} = \frac{w_{0,r+\half}+w_{0,r-\half}}{2}.
\end{equation}
The spatial derivatives of $w_0$ are two point differences:
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)_r = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.
\end{equation}
We pick the final edge states using:
\begin{equation}
\wt^{\trans}_{\ib-\half\eb_r} =
\begin{cases}
0, & (w_L \le 0 ~ {\rm AND} ~ w_R \ge 0) ~ {\rm OR} ~ |w_L + w_R| < \epsilon, \\
\wt_L, & w_L + w_R > 0, \\
\wt_R, & w_L + w_R < 0, \\
\end{cases}
\end{equation}
where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to convert $\wt^{\trans}$ to $w^{\trans}.$\\

For spherical problems, we first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.  Then we use an extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).  For example, for x-faces:

\begin{eqnarray}
\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}\right)\Delta_x\ut_{\ib-\eb_x}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib-\eb_x}\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_x}n_x, \\
\ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib}\left(\frac{\partial w_0}{\partial r}\right)_{\ib}n_x,
\end{eqnarray}
where $n_i = (\eb_r\cdot\eb_i)$.  For the y and z-faces, simply replace each instance of ``x'' with ``y'' or ``z''.
\section{Computing $\ub^{\mac}$}
In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac}$.
\section{Computing $s^{\edge}$}
\section{Computing $\ub^{\edge}$}
\end{document}

>>>>

For Cartesian problems, for the radial faces, we must include the radial derivative terms proportional to $w_0$ and $(\Ubt\cdot\eb_r) = \wt$:
\begin{eqnarray}
\wt_L &=& \wt_{\ib-\eb_r}^n + \left(\half - \frac{\dt}{h}w_{\ib-\eb_r}^n\right)\Delta_r \wt_{\ib-\eb_r}^n + \frac{\dt}{2}\wt_{\ib-\eb_r}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_r},\label{Radial utrans_L} \\
\wt_R &=& \wt_{\ib}^n + \left(\half - \frac{\dt}{h}w_{\ib}^n\right)\Delta_r \wt_{\ib}^n + \frac{\dt}{2}\wt_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib},\label{Radial utrans_R}
\end{eqnarray}
where the total radial velocity is defined as:
\begin{equation}
w_{\ib}^n = \wt_{\ib}^n + \frac{w_{0,r+\half}+w_{0,r-\half}}{2}
\end{equation}
The spatial derivatives of $w_0$ are two point differences:
\begin{equation}
\left(\frac{\partial w_0}{\partial r}\right)_{\ib} = \frac{w_{0,r+\half}-w_{0,r-\half}}{h}.
\end{equation}
We pick the final edge states using:
\begin{equation}
\wt^{\trans}_{\ib-\half\eb_r} =
\begin{cases}
0, & (w_L \le 0 ~ {\rm AND} ~ w_R \ge 0) ~ {\rm OR} ~ |w_L + w_R| < \epsilon, \\
\wt_L, & w_L + w_R > 0, \\
\wt_R, & w_L + w_R < 0, \\
\end{cases}
\end{equation}
where $w_L = \wt_L + w_{0,r-\half}$ and $w_R = \wt_R + w_{0,r-\half}$.  As mentioned before, we now call {\tt addw0} to convert $\wt^{\trans}$ to $w^{\trans}.$\\

For spherical problems, we first construct $\Ub$ by mapping $w_0$ onto a Cartesian array and adding it to $\Ubt$.  We also map $\partial w_0/\partial r$ onto a Cartesian array.  Then we use an extrapolation analogous to equations equations (\ref{Radial utrans_L}) and (\ref{Radial utrans_R}).  For example, for x-faces:

\begin{eqnarray}
\ut_L &=& \ut_{\ib-\eb_x}^n + \left(\half - \frac{\dt}{h}u_{\ib-\eb_x}^n\right)\Delta_x\ut_{\ib-\eb_x}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib-\eb_x}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib-\eb_x}(\eb_r\cdot\eb_x)_{\ib-\eb_x}, \\
\ut_R &=& \ut_{\ib}^n + \left(\half - \frac{\dt}{h}u_{\ib}^n\right)\Delta_x\ut_{\ib}^n + \frac{\dt}{2}(\Ubt\cdot\eb_r)_{\ib}^n\left(\frac{\partial w_0}{\partial r}\right)_{\ib}(\eb_r\cdot\eb_x)_{\ib}.
\end{eqnarray}
For the y and z-faces, simply replace each instance of ``x'' with ``y'' or ``z''.  Finally, we call {\tt addw0} to convert $\Ubt^{\trans}$ to $\Ub^{\trans}$.
\section{Computing $\ub^{\mac}$}
In {\tt advance\_premac.f90}, we call {\tt make\_edge\_state} to compute $\Ubt^{\mac}$.
\section{Computing $s^{\edge}$}
\section{Computing $\ub^{\edge}$}
\end{document}
